---
title: "Plotting and Statistical Analysis for Resubmission"
author: "PON"
date: "Jan/17/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Plotting and Statistic Analysis

This document contains the code we used to make plots and perform statistical testing for experiment 1-3

### 0.Prerequisite
```{r }
# import packages
# if (!require("dplyr")) install.packages("dplyr")
# if (!require("ggplot2")) install.packages("ggplot2")
# if (!require("forcats")) install.packages("forcats")
# if (!require("brms")) install.packages("brms")
# if (!require("knitr")) install.packages("knitr")
# if (!require("stringr")) install.packages("stringr")
library(dplyr)
library(ggplot2)
library(forcats)
library(brms)
library(knitr)
library(stringr)
```

### 1.Experiment 1
This section contains the code use to plot figures and perform statistical analysis for experiment1

#### 1.1 Plotting
```{r }
# import data
data_1 <- read.csv("data_experiment_1.csv")
# delete the row with NAs
data_1 <- data_1 %>% filter(!is.na(rating))
# recode the name of the variable
data_1$condition <- recode(data_1$condition, HL="High-Low", L="Low", H="High", LH="Low-High")
# change the data type of the columns
data_1$condition <- as.factor(data_1$condition)
data_1$item <- as.factor(data_1$item)
data_1$ID <- as.factor (data_1$ID)
data_1$modal <- as.factor (data_1$modal)
summary(data_1)

# reorder factor for plotting 
data_1 <-  data_1 %>% mutate(condition = fct_relevel(condition,"High","High-Low","Low-High","Low"))
summary(data_1)
```
##### 1.1.1 Plot the mean rating score across conditions
```{r }
# Plot the mean rating score across conditions
# calculate the statistics
data_mean <- data_1 %>% group_by(condition) %>% summarize(mean_rating=mean(rating),sd_rating=sd(rating),se=sd(rating) / sqrt(length(rating)))
# making a bar plot
ggplot(data=data_mean, mapping= aes(x=condition,y=mean_rating))+
  #geom_bar(stat = "identity",position = "dodge",fill="steelblue")+
  geom_col(width = 0.6,stat = "identity",position = "dodge",fill="steelblue")+
  geom_text(aes(label=round(mean_rating,1)), vjust=3, color="white", size=3.5)+
  geom_errorbar(aes(ymin=mean_rating-se, ymax=mean_rating+se),width=.5,position=position_dodge(.9))+
  labs(x = "Order of the Modals", y = expression(paste("Probability\nRating")))+
  theme(axis.text.x = element_text(color = "grey20", size = 10, angle = 0, hjust = .5, vjust = .5, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 10, angle = 0, hjust = .5, vjust = .5, face = "plain"),  
        axis.title.x = element_text(color = "grey20", size = 12, angle = 0, hjust = 0.5, vjust = -1.5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 12, angle = 0, hjust = -1.5, vjust = 0.5, face = "plain"))
ggsave("bar_plot_exp1_condition.png")

```
##### 1.1.2 Plot probability rating of different epistemic expressions
```{r }
# preparation for plotting: 
# re-code the modal column so that the high modals always precede the low modals
data_1$modal_recode <- recode(data_1$modal,may_certainly = "certainly_may", might_probably = "probably_might")
summary(data_1)
# create a column showing what modal each stimuli has
mylist <- NULL
counter <- 1
for (item in data_1$condition){
  if (item == "High") {epistemic_expression <- data_1$modal_recode[counter] %>% str_split("_") %>%
sapply("[", 1)} # split the modal_recode column by "_", and take the first element
  else if (item == "Low") {epistemic_expression <- data_1$modal_recode[counter] %>% str_split("_") %>% sapply("[", 2)} 
  else if (item == "Low-High") {epistemic_expression <- str_c(data_1$modal_recode[counter] %>% str_split("_") %>% sapply("[", 2), data_1$modal_recode[counter] %>% str_split("_") %>%
sapply("[", 1), sep = " ")} 
  else {epistemic_expression <- str_c(data_1$modal_recode[counter] %>% str_split("_") %>% sapply("[", 1), data_1$modal_recode[counter] %>% str_split("_") %>% sapply("[", 2), sep = " ")}
 
counter <- counter +1
mylist <- append(mylist, epistemic_expression)
}
data_1$epistemic_expression <- mylist
```
```{r}
# plot the probability rating of different epistemic expressions
# subset the data leaving only single model conditions
data1_sub <- data_1 %>% filter(condition == "High" | condition == "Low" )
# calculate the statistics
data1_sub_mean <- data1_sub %>% group_by(epistemic_expression) %>% summarize(mean_rating=mean(rating),sd_rating=sd(rating),se=sd(rating) / sqrt(length(rating)))
# create boxplot for data1_sub_mean
ggplot(data1_sub_mean, mapping= aes(x=fct_rev(fct_reorder(epistemic_expression,mean_rating)),
                                    y=mean_rating))+
  geom_bar(stat = "identity", position = "dodge", fill="steelblue")+
  geom_text(aes(label=round(mean_rating,1)), vjust=3, color="white", size=3.5)+
  geom_errorbar(aes(ymin=mean_rating-se, ymax=mean_rating+se),width=.5,position=position_dodge(.9))+
  labs(x = "Epistemic Modals", y = expression(paste("Probability\nRating")))+
  theme(axis.text.x = element_text(color = "grey20", size = 10, angle = 0, hjust = .5, vjust = .5, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 10, angle = 0, hjust = .5, vjust = .5, face = "plain"),  
        axis.title.x = element_text(color = "grey20", size = 12, angle = 0, hjust = 0.5, vjust = -1.5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 12, angle = 0, hjust = -0.9, vjust = 0.45, face = "plain"))
ggsave("bar_plot_exp1_modals.png")
```

#### 1.2 Statistical testing
```{r }
# Since the dependent variable, rating scores, was bounded between 0 and 100, we analyzed the data on the logit scale rather than probability scale.
# re-level condition, rescale the dependent variable (to 0~1)
data_1$condition<- factor(data_1$condition, levels = c("High-Low","Low-High","High","Low"))
data_1$rating_rescale <- data_1$rating/100

# recode the scale variable in which 1 is coded as 0.999, while 0 is coded as 0.001 (following Verkuilen, Smithson, 2012)
data_1$rating_recode <- recode(data_1$rating_rescale, `1.00` = 0.999, `0.00` = 0.001)

# change the dependent variable into logit 
data_1$rating_logit <- log(data_1$rating_recode/(1-data_1$rating_recode))

```

##### 1.2.1 maximal brm model: High-low as the baseline
```{r}
# building a linear regression model with high-low condition as the baseline
brm_max_HL_1 <- brm (rating_logit ~ 1 + condition + 
                      (1+ condition|ID) + 
                      (1+ condition|item),
                    data = data_1,
                    iter = 5000,
                    #control = list(adapt_delta = 0.99), set delta crashes the program
                    #keep it to the default which is 0.95
                    seed = 123)
summary(brm_max_HL_1)
# create a table for model output
output <- summary(brm_max_HL_1)$fixed
output <- output[,1:4]
row.names(output) <- c('Intercept', 'Low-High', 'High', 'Low')
mytable_exp1_1 <- kable(
  output,
  digits = 2,
  align = c('c','c','c','c','c'),
  format = "latex"
)

```
##### 1.2.2 maximal brm model: Low-High as the baseline
```{r }
# building a linear regression model with Low-High condition as the baseline
data_1$condition<- factor(data_1$condition, levels = c("Low-High","High-Low","High","Low"))
brm_max_LH_1 <- brm (rating_logit ~ 1 + condition + 
                      (1+ condition|ID) + 
                      (1+ condition|item),
                    data = data_1,
                    iter = 5000,
                    control = list(adapt_delta = 0.99),
                    seed = 123)
summary(brm_max_LH_1)
# create a table for model output
output <- summary(brm_max_LH_1)$fixed
output <- output[,1:4]
row.names(output) <- c('Intercept', 'High-Low', 'High', 'Low')
mytable_exp1_2 <- kable(
  output,
  digits = 2,
  align = c('c','c','c','c','c'),
  format = "latex"
)

```

##### 1.2.3 Bayes factor analysis
The maximal model used in the previous section reveals no statistically meaningful difference between "High-Low" and "Low-High" condition. This section further investigates the likelihood ratio of H0 over H1.
```{r}
# Subset the data_1 keeping only the two nested modal conditions
data_1_sub <- data_1 %>% filter(data_1$condition == "High-Low" | data_1$condition == "Low-High")
# create a list of uninformative prior for slope
# the prior has a mean of -0.2 (the estimate from the previous model) and 0
# the sd of the prior varies but not smaller than 1, thus not too informative
prior_list_bf = c('normal(0, 10)', 'normal(0, 5)','normal(0, 1)','normal(-0.2, 10)', 'normal(-0.2, 5)','normal(-0.2, 1)')
# create an empty data frame to store the output of bf analysis
tab_bf_analysis <- data.frame(prior = character(),
                     estimate = double(),
                     bf01 = double())
# loops over a variety of priors for the slope of condition variable in brm model
for (item in prior_list_bf){
  priors_mixed = c(set_prior(item, coef = 'conditionLowMHigh'))
  brm_max_data_1_subset<- brm (rating_logit ~ 1 + condition 
                            + (1+ condition|ID)
                            + (1+ condition|item),
                    data = data_1_sub,
                    prior = priors_mixed, 
                    sample_prior = TRUE,
                    iter = 5000,
                    seed = 123)
  bf <- hypothesis(brm_max_data_1_subset,hypothesis = "conditionLowMHigh = 0")
  p <- item
  e <- bf$hypothesis[[2]]
  b <- bf$hypothesis[[6]]
  new_row <- c(p,e,b)
  tab_bf_analysis[nrow(tab_bf_analysis)+1,] = new_row
}
write.csv(tab_bf_analysis, "tab_bf_analysis.csv")
print(tab_bf_analysis)
# The result reveals moderate to strong evidence for H0, which says
# the coefficient of condition is zero
# create latex table
tab_bf_analysis$estimate <- as.double(tab_bf_analysis$estimate)
tab_bf_analysis$bf01 <- as.double(tab_bf_analysis$bf01 )
mytable_exp1_3 <- kable(
  tab_bf_analysis,
  digits = 2,
  align = c('c','c','c'),
  format = "latex"
)
write.csv(tab_bf_analysis, "bf_analysis_exp1_1.csv")
```
##### 1.2.4 sensitivity analysis of prior 
This section loops over a variety of priors for the slope of condition variable and see how they influence bf01
```{r}
tab_sa <- data.frame(prior = character(),
                     estimate = double(),
                     bf01 = double())
# select a set of means for the prior
a <- seq(-0.2,0.2,by=0.1)
# select a set of sds for the prior
b <- seq(0.1,1,by=0.1)
# cross a and b to get a set of mean and sd combinations
c <- crossing(a,b)
# create a list of priors containing all mean and sd combinations
prior_list <- c()
for (row in 1:nrow(c)){
  first <- as.character(c[row,1])
  second <- as.character(c[row,2])
  label <- paste0("normal","(",first,",",second,")")
prior_list <- append(prior_list, label)
}
# loops over a variety of priors for the slope of condition variable in brm model
for (item in prior_list){
  priors_mixed = c(set_prior(item, coef = 'conditionLowMHigh'))
  brm_max_data_1_subset<- brm (rating_logit ~ 1 + condition 
                            + (1+ condition|ID)
                            + (1+ condition|item),
                    data = data_1_sub,
                    prior = priors_mixed, 
                    sample_prior = TRUE,
                    iter = 5000,
                    seed = 123)
  bf <- hypothesis(brm_max_data_1_subset,hypothesis = "conditionLowMHigh = 0")
  p <- item
  e <- bf$hypothesis[[2]]
  b <- bf$hypothesis[[6]]
  new_row <- c(p,e,b)
  tab_sa[nrow(tab_sa)+1,] = new_row
}
write.csv(tab_sa, "bf_analysis_exp1_2.csv")
# calculate the proportion of BFs in favor of H0
print(sum(tab_sa[,3]>=1)/length(tab_sa[,3]))
```

#### 1.3 An additional question in experiment1
Was the rating score of the modal "probably" significantly different from the rating score of the modal "possibly"? The following code used linear mixed effects modal to explore this question
```{r }
# subset data_1, leaving only rating scores for modal "probably and possibly"
data_1$epistemic_expression <- as.factor (data_1$epistemic_expression)
data1_sub2 <- data_1 %>% filter(epistemic_expression == "probably" | epistemic_expression == "possibly")
# build a linear mixed effect modal using epistemic expression to predict rating score
brm_max_exp1_expression <- brm (rating_logit ~ 1 + epistemic_expression + 
                      (1+ epistemic_expression|ID) + 
                      (1+ epistemic_expression|item),
                    data = data1_sub2,
                    iter = 5000,
                    control = list(adapt_delta = 0.99),
                    seed = 123)
summary(brm_max_exp1_expression)

```


### 2.Experiment 2
This section contains the code use to plot figures and perform statistical analysis for experiment2

#### 2.1 Plotting
```{r }
# import data
data_2 <- read.csv("data_experiment_2.csv")
# delete the row with NAs
data_2 <- data_2 %>% filter(!is.na(rating))
# recode the name of the variable
data_2$condition <- recode(data_2$condition, HL="High-Low", L="Low", H="High", LH="Low-High")
# change the data type of the columns
data_2$condition <- as.factor(data_2$condition)
data_2$item <- as.factor(data_2$item)
data_2$ID <- as.factor (data_2$ID)
data_2$modal <- as.factor (data_2$modal)
summary(data_2)

# reorder factor for plotting 
data_2 <-  data_2 %>% mutate(condition = fct_relevel(condition,"High","High-Low","Low-High","Low"))
summary(data_2)

```
##### 2.1.1 Plot the mean rating score across conditions
```{r }
# Plot the mean rating score across conditions
# calculate the statistics
data_mean_2 <- data_2 %>% group_by(condition) %>% summarize(mean_rating=mean(rating),sd_rating=sd(rating),se=sd(rating) / sqrt(length(rating)))
# making a bar plot
ggplot(data=data_mean_2, mapping= aes(x=condition,y=mean_rating))+
  #geom_bar(stat = "identity",position = "dodge",fill="steelblue")+
  geom_col(width = 0.6,stat = "identity",position = "dodge",fill="steelblue")+
  geom_text(aes(label=round(mean_rating,1)), vjust=3, color="white", size=3.5)+
  geom_errorbar(aes(ymin=mean_rating-se, ymax=mean_rating+se),width=.5,position=position_dodge(.9))+
  labs(x = "Order of the Modals", y = expression(paste("Probability\nRating")))+
  theme(axis.text.x = element_text(color = "grey20", size = 10, angle = 0, hjust = .5, vjust = .5, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 10, angle = 0, hjust = .5, vjust = .5, face = "plain"),  
        axis.title.x = element_text(color = "grey20", size = 12, angle = 0, hjust = 0.5, vjust = -1.5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 12, angle = 0, hjust = -1.5, vjust = 0.55, face = "plain"))
ggsave("bar_plot_exp2_condition.png")

```
##### 2.1.2 Plot the mean rating score across different epistemic expressions
```{r }
# preparation for plotting: 
# re-code the modal column so that the high modals always precede the low modals
data_2$modal_recode <- recode(data_2$modal,may_cert= "certainly_may", may_def = "definitely_may", might_cert = "certainly_might", might_def = "definitely_might")
summary(data_2)
# create a column showing what modal each stimuli has
mylist <- NULL
counter <- 1
for (item in data_2$condition){
  if (item == "High") {epistemic_expression <- data_2$modal_recode[counter] %>% str_split("_") %>%
sapply("[", 1)} # split the modal_recode column by "_", and take the first element
  else if (item == "Low") {epistemic_expression <- data_2$modal_recode[counter] %>% str_split("_") %>% sapply("[", 2)} 
  else if (item == "Low-High") {epistemic_expression <- str_c(data_2$modal_recode[counter] %>% str_split("_") %>% sapply("[", 2), data_2$modal_recode[counter] %>% str_split("_") %>%
sapply("[", 1), sep = " ")} 
  else {epistemic_expression <- str_c(data_2$modal_recode[counter] %>% str_split("_") %>% sapply("[", 1), data_2$modal_recode[counter] %>% str_split("_") %>% sapply("[", 2), sep = " ")}
 
counter <- counter +1
mylist <- append(mylist, epistemic_expression)
}
data_2$epistemic_expression <- mylist
```
```{r}
# plot the probability rating of different epistemic expressions
# subset the data leaving only single model conditions
data2_sub <- data_2 %>% filter(condition == "High" | condition == "Low" )
# calculate the statistics
data2_sub_mean <- data2_sub %>% group_by(epistemic_expression) %>% summarize(mean_rating=mean(rating),sd_rating=sd(rating),se=sd(rating) / sqrt(length(rating)))
# create boxplot for data1_sub_mean
ggplot(data2_sub_mean, mapping= aes(x=fct_rev(fct_reorder(epistemic_expression,mean_rating)),
                                    y=mean_rating))+
  #geom_bar(stat = "identity", position = "dodge", fill="steelblue")+
  geom_col(width = 0.6,stat = "identity",position = "dodge",fill="steelblue")+
  geom_text(aes(label=round(mean_rating,1)), vjust=3, color="white", size=3.5)+
  geom_errorbar(aes(ymin=mean_rating-se, ymax=mean_rating+se),width=.5,position=position_dodge(.9))+
  labs(x = "Epistemic Modals", y = expression(paste("Probability\nRating")))+
  theme(axis.text.x = element_text(color = "grey20", size = 10, angle = 0, hjust = .5, vjust = .5, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 10, angle = 0, hjust = .5, vjust = .5, face = "plain"),  
        axis.title.x = element_text(color = "grey20", size = 12, angle = 0, hjust = 0.5, vjust = -1.5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 12, angle = 0, hjust = -1.5, vjust = 0.55, face = "plain"))
ggsave("bar_plot_exp2_modals.png")
```

#### 2.2 Statistical testing
```{r }
# Since the dependent variable, rating scores, was bounded between 0 and 100, we analyzed the data on the logit scale rather than probability scale.
# re-level condition, rescale the dependent variable (to 0~1)
data_2$condition<- factor(data_2$condition, levels = c("High-Low","Low-High","High","Low"))
data_2$rating_rescale <- data_2$rating/100

# recode the scale variable in which 1 is coded as 0.999, while 0 is coded as 0.001 (following Verkuilen, Smithson, 2012)
data_2$rating_recode <- recode(data_2$rating_rescale, `1.00` = 0.999, `0.00` = 0.001)

# change the dependent variable into logit 
data_2$rating_logit <- log(data_2$rating_recode/(1-data_2$rating_recode))

```
##### 2.2.1 maximal brm model: High-low as the baseline
```{r}
# building a linear regression model with high-low condition as the baseline
brm_max_HL_2 <- brm (rating_logit ~ 1 + condition + 
                      (1+ condition|ID) + 
                      (1+ condition|item),
                    data = data_2,
                    iter = 5000,
                    #control = list(adapt_delta = 0.99),
                    seed = 123)
summary(brm_max_HL_2)
# create a table for model output
output <- summary(brm_max_HL_2)$fixed
output <- output[,1:4]
row.names(output) <- c('Intercept', 'Low-High', 'High', 'Low')
mytable_exp2_1 <- kable(
  output,
  digits = 2,
  align = c('c','c','c','c','c'),
  format = "latex"
)
```
##### 2.2.2 maximal brm model: Low-High as the baseline
```{r }
# building a linear regression model with Low-High condition as the baseline
data_2$condition<- factor(data_2$condition, levels = c("Low-High","High-Low","High","Low"))
brm_max_LH_2 <- brm (rating_logit ~ 1 + condition + 
                      (1+ condition|ID) + 
                      (1+ condition|item),
                    data = data_2,
                    iter = 5000,
                    control = list(adapt_delta = 0.99),
                    seed = 123)
summary(brm_max_LH_2)
# create a table for model output
output <- summary(brm_max_LH_2)$fixed
output <- output[,1:4]
row.names(output) <- c('Intercept', 'High-Low', 'High', 'Low')
mytable_exp2_2 <- kable(
  output,
  digits = 2,
  align = c('c','c','c','c','c'),
  format = "latex"
)

```
##### 2.2.3 Bayes factor analysis
The maximal model used in the previous section reveals no statistically meaningful difference between "High-Low" and "Low-High" condition. This section further investigates the likelihood ratio of H0 over H1.
```{r}
# Subset the data_1 keeping only the two nested modal conditions
data_2_sub <- data_2 %>% filter(data_2$condition == "High-Low" | data_2$condition == "Low-High")
# create a list of uninformative prior for slope
# the prior has a mean of -0.12 (the estimate from the previous model) and 0
# the sd of the prior varies but not smaller than 1, thus not too informative
prior_list_bf_2 = c('normal(0, 10)', 'normal(0, 5)','normal(0, 1)','normal(-0.12, 10)', 'normal(-0.12, 5)','normal(-0.12, 1)')
# create an empty data frame to store the output of bf analysis
tab_bf_analysis_2 <- data.frame(prior = character(),
                     estimate = double(),
                     bf01 = double())
# loops over a variety of priors for the slope of condition variable in brm model
for (item in prior_list_bf_2){
  priors_mixed = c(set_prior(item, coef = 'conditionLowMHigh'))
  brm_max_data_2_subset<- brm (rating_logit ~ 1 + condition 
                            + (1+ condition|ID)
                            + (1+ condition|item),
                    data = data_2_sub,
                    prior = priors_mixed, 
                    sample_prior = TRUE,
                    iter = 5000,
                    seed = 123)
  bf <- hypothesis(brm_max_data_2_subset,hypothesis = "conditionLowMHigh = 0")
  p <- item
  e <- bf$hypothesis[[2]]
  b <- bf$hypothesis[[6]]
  new_row <- c(p,e,b)
  tab_bf_analysis_2[nrow(tab_bf_analysis_2)+1,] = new_row
}
write.csv(tab_bf_analysis_2, "tab_bf_analysis_2.csv")
print(tab_bf_analysis_2)
# The result reveals moderate to strong evidence for H0, which says
# the coefficient of condition is zero
# create latex table
tab_bf_analysis_2$estimate <- as.double(tab_bf_analysis_2$estimate)
tab_bf_analysis_2$bf01 <- as.double(tab_bf_analysis_2$bf01 )
mytable_exp2_3 <- kable(
  tab_bf_analysis_2,
  digits = 2,
  align = c('c','c','c'),
  format = "latex"
)
write.csv(tab_bf_analysis_2, "bf_analysis_exp2_1.csv")
```
##### 2.2.4 sensitivity analysis of prior 
This section loops over a variety of priors for the slope of condition variable and see how they influence bf01
```{r}
tab_sa_2 <- data.frame(prior = character(),
                     estimate = double(),
                     bf01 = double())
# select a set of means for the prior
a <- seq(-0.12,0.12,by=0.06)
# select a set of sds for the prior
b <- seq(0.1,1,by=0.1)
# cross a and b to get a set of mean and sd combinations
c <- crossing(a,b)
# create a list of priors containing all mean and sd combinations
prior_list <- c()
for (row in 1:nrow(c)){
  first <- as.character(c[row,1])
  second <- as.character(c[row,2])
  label <- paste0("normal","(",first,",",second,")")
prior_list <- append(prior_list, label)
}
# loops over a variety of priors for the slope of condition variable in brm model
for (item in prior_list){
  priors_mixed = c(set_prior(item, coef = 'conditionLowMHigh'))
  brm_max_data_2_subset<- brm (rating_logit ~ 1 + condition 
                            + (1+ condition|ID)
                            + (1+ condition|item),
                    data = data_2_sub,
                    prior = priors_mixed, 
                    sample_prior = TRUE,
                    iter = 5000,
                    seed = 123)
  bf <- hypothesis(brm_max_data_2_subset,hypothesis = "conditionLowMHigh = 0")
  p <- item
  e <- bf$hypothesis[[2]]
  b <- bf$hypothesis[[6]]
  new_row <- c(p,e,b)
  tab_sa_2[nrow(tab_sa_2)+1,] = new_row
}
write.csv(tab_sa_2, "bf_analysis_exp2_2.csv")
# calculate the proportion of BFs in favor of H0
print(sum(tab_sa_2[,3]>=1)/length(tab_sa_2[,3]))
```




### Experiment 3
This section contains the code use to plot figures and perform statistical analysis for experiment3

#### 3.1 Plotting

```{r }
# import data
data_3 <- read.csv("data_experiment_3.csv")
# delete the row with NAs
data_3 <- data_3 %>% filter(!is.na(rating))
# recode the name of the variable
data_3$condition <- recode(data_3$condition, HL="High-Low", L="Low", H="High", LH="Low-High")
# change the data type of the columns
data_3$condition <- as.factor(data_3$condition)
data_3$item <- as.factor(data_3$item)
data_3$ID <- as.factor (data_3$ID)
data_3$modal <- as.factor (data_3$modal)
summary(data_3)

# reorder factor for plotting 
data_3 <-  data_3 %>% mutate(condition = fct_relevel(condition,"High","High-Low","Low-High","Low"))
summary(data_3)
```

##### 3.1.1 Plot the mean rating score across conditions
```{r }
# Plot the mean rating score across conditions
# calculate the statistics
data_mean_3 <- data_3 %>% group_by(condition) %>% summarize(mean_rating=mean(rating),sd_rating=sd(rating),se=sd(rating) / sqrt(length(rating)))
# making a bar plot
ggplot(data=data_mean_3, mapping= aes(x=condition,y=mean_rating))+
  #geom_bar(stat = "identity",position = "dodge",fill="steelblue")+
  geom_col(width = 0.6,stat = "identity",position = "dodge",fill="steelblue")+
  geom_text(aes(label=round(mean_rating,1)), vjust=3, color="white", size=3.5)+
  geom_errorbar(aes(ymin=mean_rating-se, ymax=mean_rating+se),width=.5,position=position_dodge(.9))+
  labs(x = "Order of the Modals", y = expression(paste("Probability\nRating")))+
  theme(axis.text.x = element_text(color = "grey20", size = 10, angle = 0, hjust = .5, vjust = .5, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 10, angle = 0, hjust = .5, vjust = .5, face = "plain"),  
        axis.title.x = element_text(color = "grey20", size = 12, angle = 0, hjust = 0.5, vjust = -1.5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 12, angle = 0, hjust = -1.5, vjust = 0.5, face = "plain"))
ggsave("bar_plot_exp3_condition.png")
```

##### 3.1.2 Plot the mean rating score across conditions across modal combinations
```{r }
# Plot the mean rating score across conditions across modal combinations
data_3 <-  data_3 %>% mutate(condition = fct_relevel(condition,"High","High-Low","Low-High","Low"))
data_mean_3_2 <- data_3 %>% group_by(condition,modal) %>% summarize(mean_rating=mean(rating),sd_rating=sd(rating),se=sd(rating) / sqrt(length(rating)))
# making a bar plot
ggplot(data=data_mean_3_2, mapping= aes(x= modal,y=mean_rating, fill = condition))+
  geom_bar(stat = "identity",position = "dodge")+
  #geom_col(width = 0.7,stat = "identity",position = "dodge",fill="steelblue")+
  #geom_text(aes(label=round(mean_rating,2)), vjust=3, color="white", size=3.5)+
  geom_errorbar(aes(ymin=mean_rating-se, ymax=mean_rating+se),width=.5,position=position_dodge(.9))+
  #labs(x = "Order of the Modals", y = expression(paste("Probability\nRating")))+
  theme(axis.text.x = element_text(color = "grey20", size = 10, angle = 45, hjust = .5, vjust = .5, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 10, angle = 0, hjust = .5, vjust = .5, face = "plain"),  
        axis.title.x = element_text(color = "grey20", size = 12, angle = 0, hjust = 0.5, vjust = -1.5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 12, angle = 0, hjust = -1.5, vjust = 0.5, face = "plain"))
#ggsave("bar_plot_exp3_condition.png")


```
##### 3.1.3 Plot the mean rating score across different modals

```{r }
# preparation for plotting: 
# create a column showing what modal each stimuli has
mylist <- NULL
counter <- 1
for (item in data_3$condition){
  if (item == "High") {epistemic_expression <- data_3$modal[counter] %>% str_split("_") %>%
sapply("[", 1)} # split the modal_recode column by "_", and take the first element
  else if (item == "Low") {epistemic_expression <- data_3$modal[counter] %>% str_split("_") %>% sapply("[", 2)} 
  else if (item == "Low-High") {epistemic_expression <- str_c(data_3$modal[counter] %>% str_split("_") %>% sapply("[", 2), data_3$modal[counter] %>% str_split("_") %>%
sapply("[", 1), sep = " ")} 
  else {epistemic_expression <- str_c(data_3$modal[counter] %>% str_split("_") %>% sapply("[", 1), data_3$modal[counter] %>% str_split("_") %>% sapply("[", 2), sep = " ")}
 
counter <- counter +1
mylist <- append(mylist, epistemic_expression)
}
data_3$epistemic_expression <- mylist
data_3$epistemic_expression <- as.factor (data_3$epistemic_expression)

```

```{r }
# plot the probability rating of different epistemic expressions
# subset the data leaving only single model conditions
data3_sub <- data_3 %>% filter(condition == "High" | condition == "Low" )
# calculate the statistics
data3_sub_mean <- data3_sub %>% group_by(epistemic_expression) %>% summarize(mean_rating=mean(rating),sd_rating=sd(rating),se=sd(rating) / sqrt(length(rating)))
# create boxplot for data1_sub_mean
ggplot(data3_sub_mean, mapping= aes(x=fct_rev(fct_reorder(epistemic_expression,mean_rating)),
                                    y=mean_rating))+
  geom_bar(stat = "identity", position = "dodge", fill="steelblue")+
  geom_text(aes(label=round(mean_rating,1)), vjust=3, color="white", size=3.5)+
  geom_errorbar(aes(ymin=mean_rating-se, ymax=mean_rating+se),width=.5,position=position_dodge(.9))+
  labs(x = "Epistemic Modals", y = expression(paste("Probability\nRating")))+
  theme(axis.text.x = element_text(color = "grey20", size = 10, angle = 0, hjust = .5, vjust = .5, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 10, angle = 0, hjust = .5, vjust = .5, face = "plain"),  
        axis.title.x = element_text(color = "grey20", size = 12, angle = 0, hjust = 0.5, vjust = -1.5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 12, angle = 0, hjust = -1.5, vjust = 0.57, face = "plain"))
ggsave("bar_plot_exp3_modals.png")
```

#### 3.2 Statistical testing
```{r }
# Since the dependent variable, rating scores, was bounded between 0 and 100, we analyzed the data on the logit scale rather than probability scale.
# re-level condition, rescale the dependent variable (to 0~1)
data_3$condition<- factor(data_3$condition, levels = c("High-Low","Low-High","High","Low"))
data_3$rating_rescale <- data_3$rating/100

# recode the scale variable in which 1 is coded as 0.999, while 0 is coded as 0.001 (following Verkuilen, Smithson, 2012)
data_3$rating_recode <- recode(data_3$rating_rescale, `1.00` = 0.999, `0.00` = 0.001)

# change the dependent variable into logit 
data_3$rating_logit <- log(data_3$rating_recode/(1-data_3$rating_recode))
```

##### 3.2.1 maximal brm model: High-low as the baseline
```{r }
# building a linear regression model with high-low condition as the baseline
brm_max_HL_3 <- brm (rating_logit ~ 1 + condition + 
                      (1+ condition|ID) + 
                      (1+ condition|item),
                    data = data_3,
                    iter = 5000,
                    control = list(adapt_delta = 0.99),
                    seed = 123)
summary(brm_max_HL_3)
# create a table for model output
output <- summary(brm_max_HL_3)$fixed
output <- output[,1:4]
row.names(output) <- c('Intercept', 'Low-High', 'High', 'Low')
mytable_exp3_1 <- kable(
  output,
  digits = 2,
  align = c('c','c','c','c','c'),
  format = "latex"
)
```

##### 3.2.2 maximal brm model: Low-High as the baseline
```{r }
# building a linear regression model with Low-High condition as the baseline
data_3$condition<- factor(data_3$condition, levels = c("Low-High","High-Low","High","Low"))
brm_max_LH_3 <- brm (rating_logit ~ 1 + condition + 
                      (1+ condition|ID) + 
                      (1+ condition|item),
                    data = data_3,
                    iter = 5000,
                    control = list(adapt_delta = 0.99),
                    seed = 123)
summary(brm_max_LH_3)
# create a table for model output
output <- summary(brm_max_LH_3)$fixed
output <- output[,1:4]
row.names(output) <- c('Intercept', 'High-Low', 'High', 'Low')
mytable_exp3_2 <- kable(
  output,
  digits = 2,
  align = c('c','c','c','c','c'),
  format = "latex"
)

```

##### 3.2.3 Bayes factor analysis
The maximal model used in the previous section reveals no statistically meaningful difference between "High-Low" and "Low-High" condition. This section further investigates the likelihood ratio of H0 over H1.
```{r}
# Subset the data_1 keeping only the two nested modal conditions
data_3_sub <- data_3 %>% filter(data_3$condition == "High-Low" | data_3$condition == "Low-High")
# create a list of uninformative prior for slope
# the prior has a mean of 0.01 (the estimate from the previous model)
# the sd of the prior varies but not smaller than 1, thus not too informative
prior_list_bf_3 = c('normal(0, 10)', 'normal(0, 5)','normal(0, 1)','normal(0.01, 10)', 'normal(0.01, 5)','normal(0.01, 1)')
# create an empty data frame to store the output of bf analysis
tab_bf_analysis_3 <- data.frame(prior = character(),
                     estimate = double(),
                     bf01 = double())
# loops over a variety of priors for the slope of condition variable in brm model
for (item in prior_list_bf_3){
  priors_mixed = c(set_prior(item, coef = 'conditionLowMHigh'))
  brm_max_data_3_subset<- brm (rating_logit ~ 1 + condition 
                            + (1+ condition|ID)
                            + (1+ condition|item),
                    data = data_3_sub,
                    prior = priors_mixed, 
                    sample_prior = TRUE,
                    iter = 5000,
                    seed = 123)
  bf <- hypothesis(brm_max_data_3_subset,hypothesis = "conditionLowMHigh = 0")
  p <- item
  e <- bf$hypothesis[[2]]
  b <- bf$hypothesis[[6]]
  new_row <- c(p,e,b)
  tab_bf_analysis_3[nrow(tab_bf_analysis_3)+1,] = new_row}

write.csv(tab_bf_analysis_3, "tab_bf_analysis_3.csv")
print(tab_bf_analysis_3)
# The result reveals moderate to strong evidence for H0, which says
# the coefficient of condition is zero
# create latex table
tab_bf_analysis_3$estimate <- as.double(tab_bf_analysis_3$estimate)
tab_bf_analysis_3$bf01 <- as.double(tab_bf_analysis_3$bf01 )
mytable_exp3_3 <- kable(
  tab_bf_analysis_3,
  digits = 2,
  align = c('c','c','c'),
  format = "latex"
)
write.csv(tab_bf_analysis_3, "bf_analysis_exp3_1.csv")
```
##### 3.2.4 sensitivity analysis of prior 
This section loops over a variety of priors for the slope of condition variable and see how they influence bf01
```{r}
tab_sa_3 <- data.frame(prior = character(),
                     estimate = double(),
                     bf01 = double())
# select a set of means for the prior
a <- seq(-0.01,0.01,by=0.005)
# select a set of sds for the prior
b <- seq(0.1,1,by=0.1)
# cross a and b to get a set of mean and sd combinations
c <- crossing(a,b)
# create a list of priors containing all mean and sd combinations
prior_list <- c()
for (row in 1:nrow(c)){
  first <- as.character(c[row,1])
  second <- as.character(c[row,2])
  label <- paste0("normal","(",first,",",second,")")
prior_list <- append(prior_list, label)
}
# loops over a variety of priors for the slope of condition variable in brm model
for (item in prior_list){
  priors_mixed = c(set_prior(item, coef = 'conditionLowMHigh'))
  brm_max_data_3_subset<- brm (rating_logit ~ 1 + condition 
                            + (1+ condition|ID)
                            + (1+ condition|item),
                    data = data_3_sub,
                    prior = priors_mixed, 
                    sample_prior = TRUE,
                    iter = 5000,
                    seed = 123)
  bf <- hypothesis(brm_max_data_3_subset,hypothesis = "conditionLowMHigh = 0")
  p <- item
  e <- bf$hypothesis[[2]]
  b <- bf$hypothesis[[6]]
  new_row <- c(p,e,b)
  tab_sa_3[nrow(tab_sa_3)+1,] = new_row
}
write.csv(tab_sa_3, "bf_analysis_exp3_2.csv")
# calculate the proportion of BFs in favor of H0
print(sum(tab_sa_3[,3]>=1)/length(tab_sa_3[,3]))
```